Synthesizing Symmetric Lenses

### 2018-07-01 papers/SymmetricLenses/main.tex
minor changes

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amsmath}
\usepackage{nccmath}
\usepackage{mathtools}
\usepackage{bussproofs}
\usepackage{varwidth}
\usepackage{amsthm}
\usepackage{csvsimple}
\usepackage{thmtools,thm-restate}
\usepackage{changepage}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage[inline]{enumitem}
\usepackage{multirow,bigdelim}
\usepackage{multicol}
\usepackage{siunitx}
\usepackage{listings}
\usepackage{letltxmacro}
\usepackage{sansmath}
\usepackage{url}
\usepackage{flushend}
\usepackage{microtype}
\usepackage[utf8]{inputenc}
\usepackage{mathpartir}
\usepackage{empheq}
\usepackage{array}
\usepackage{pgfplots}
\usepackage{stmaryrd}
\usepackage{courier}
\usepackage{qtree}
\usepackage[normalem]{ulem}
\usepackage{relsize}
\usepackage{tikz}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{tabularx}
\usepackage{stackengine}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage{remreset}
\usepackage{tabulary}
\usepackage{xspace}

@@

\begin{figure}
  % https://tex.stackexchange.com/questions/168741/showing-two-listings-in-a-table-side-by-side
  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}[
stringstyle=\upshape\sffamily
]
Jane Doe: 38000
John Public: 37500
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingA}{\box0}

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}[
stringstyle=\upshape\sffamily
]
Name,Company
Jane Doe,Healthcare Inc.
John Public,Insurance Co.
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingB}{\box0}

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}
let salary = number | "N/A"
let emp_salary = name . " " . name . ": " salary
let emp_salaries = "" | emp_salary . ("\n" emp_salary)*
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingC}{\box0}

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}
let company = (name . ("Co." | "Inc." | "Ltd.")) | "UNK" in
let emp_ins = name . " " . name "," company in
let header = "Name,Company" in
let emp_insurance = header . ("\n" . emp_ins)*
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingD}{\box0}

  \centering
  \resizebox{\columnwidth}{!}{%
    \begin{tabular}{|>{\columncolor{lightyellow}}c|>{\columncolor{lightgreen}}c|}
      \hline
      \texttt{company.mgmt} data file & \texttt{company.hr} data file \\
      \hline
      \listingA & \listingB \\
      \hline
      \cellcolor{vlightyellow}\texttt{company.mgmt} regular expression & 
      \cellcolor{vlightgreen}\texttt{company.hr}  regular expression\\
      \hline
      \cellcolor{vlightyellow}\listingC & \cellcolor{vlightgreen}\listingD \\
      \hline
    \end{tabular}
  }
  \caption{Example data files and corresponding regular expressions
  used by management and HR, respectively, to represent employee salaries and health
  insurance companies.  }  
  \label{fig:minimized-representations}
\end{figure}

In this section, we give an overview of simple symmetric lenses and
our corresponding synthesis algorithm using an example drawn from a
hypothetical U.S. company.  In this company, management and
human resources (HR) store information about employees in separate files:
management records the names of employees and their salaries while
HR stores the names of employees and their health
insurance providers.  Figure~\ref{fig:minimized-representations} gives
examples of the two types of files and regular expressions describing
the two formats.
The company uses simple symmetric lenses to keep these files synchronized.
If management adds a new employee, ``Chris Roe: 32500'',  a lens adds 
the corresponding entry ``Chris Roe, UNK'' to HR's file.  The
sentinel value ``UNK'' represents the fact that the employee's
insurance company is currently unknown.  A similar update happens if
HR adds a new employee before management, in which case the sentinel
value ``N/A'' reflects the unknown salary information. 
Furthermore, if HR
corrects an error in an employee's name, say changing ``John Public''
to ``Jon Public'', the lens mirrors that name change in management's
file. Not all the data is mirrored, however. The management file 
is not updated in response to insurance changes, and the HR file
is oblivious to salary changes.  Simple symmetric lenses are
appropriate for keeping these two files synchronized because they have
a mix of shared and unshared information.

@@

\subsection{Simple Symmetric Lenses}
A simple symmetric lens between formats described by types $\Regex$
and $\RegexAlt$ 
is defined by four functions: 
\begin{enumerate}
\item $\CreateR \OfType \Regex \rightarrow \RegexAlt$
\item $\CreateL \OfType \RegexAlt \rightarrow \Regex$
\item $\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$
\item $\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$
\end{enumerate}
\noindent
that collectively satisfy four ``round-tripping'' laws: 
\begin{gather}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{s})}{s} = s\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{t})}{t} = t\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{s}{t})}{s} = s\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{t}{s})}{t} = t
\end{gather}
\noindent
The two \emph{create} functions provide default values for times when
data appears on one side but not the other (eg, the ``N/A'' salary
entry added to the management file when HR inserts a new employee).
The two \emph{put} functions propagate edits from one format to the
other by combining the change from one file with the previous value
from the other file. The roundtripping laws guarantee that identify edits to one file do
not change the data in the other file.

%%For synchronization, when the original company.hr or company.mgmt
%%file is created, a synchronized file of the other form is created
%%from it with $\CreateR$ or $\CreateL$ (depending on which file was
%%initially created). On future edits, changes are made with the puts,
%%a \PutR would propagate the shared information from the company.hr
%%file, and knit it together with the information specific to the company.mgmt file.

We use regular expressions to denote the types of data formats, and we
impose a typing discipline on our lenses.  Specifically, if 
$\Regex$ and $\RegexAlt$ are regular expressions, then the notation
\lstinline{$\Lens$ : $\Regex \Leftrightarrow \RegexAlt$} indicates that $\Lens$
is a simple symmetric lens between $\LanguageOf{\Regex}$ and
$\LanguageOf{\RegexAlt}$. 
When clear from context, we use the term \textit{lens} to denote a
simple symmetric lens.

\subsection{Simple Symmetric Lens Combinators} 
\ksf{This section needs more polishing}
Generally, we do not expect programmers to define simple symmetric lenses by
writing the four functions by hand and showing they satisfy the
roundtripping laws.  Instead, we provide simple symmetric lens
combinators to build up the requisite lenses from simpler ones.
We explain these combinators by using them to define a lens between
the two employee data formats in Figure~\ref{fig:minimized-representations}. 
The finished lens, which appears in Figure~\ref{fig:example_lens}, has
the type 
\lstinline{emp_salaries} $\Leftrightarrow$ \lstinline{emp_insurance}, where
\lstinline{emp_salaries} and \lstinline{emp_insurance} are regular
expressions defined in Figure~\ref{fig:minimized-representations}.

The simplest combinator is the identity lens, which takes as an argument a
regular expression, and performs the 
identity function on strings in that regular expression. The identity lens fully
matches the data formats: when one data format is updated, the other format gets
the exact same value. In our example, when a name is updated, it is also
expected to be updated in the other format. The identity lens propagates that
information.
%
\begin{lstlisting}
id(name) : name $\Leftrightarrow$ name
\end{lstlisting}
%

In contrast to the identity lens, the disconnect lens leaves data
unaligned. The disconnect lens takes four arguments: two regular 
expressions and two strings. The regular expressions specify the
formats of the data on the two sides while the strings provide default
values.  In our extended example, salaries and
insurance companies unrelated, so we use a disconnect lens.  The
defaults are strings that indicate the information is missing.
%
\begin{lstlisting}
disc(salary,company,"N/A","UNK") : salary $\Leftrightarrow$ company
\end{lstlisting}
%
When building a disconnect lens between two strings, one of which is merely the
empty string, the lenses \lstinline{ins} and \lstinline{del} can instead be
used, where \lstinline{ins($t$) = disc("",$t$,"",$t$)} and
\lstinline{del($s$) = disc($s$,"",$s$,"")}.

Finally, there are a number of lenses that compositionally build up new lenses
from old ones. 
For example, data can be concatenated, and so we provide
the \lstinline{concat} combinator that sequences lenses, usually
written using an infix notation \lstinline{$\Lens_1$ . $\Lens_2$}:
%
\begin{lstlisting}
id(name) . ins(",") . id(name) . del(": ") . ins(",")
\end{lstlisting}
% concat(ins(header),employees_lens) : emp_salaries $\Leftrightarrow$ emp_insurance
%

Data can also appear in multiple forms. For example,
a list of employees with insurance information can either be empty or
be a sequence of employee/insurance records: \ksf{This example is
vacuous as you could eliminate the or in favor of the *.  Do somethign
else to introduce or?}
%
\begin{lstlisting}
or(id(""),employees_lens) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_ins) . ("\n" emp_ins)*
\end{lstlisting}

Lastly, the \IterateLens lens synchronizes repeated data, aligning the number of
times the data is repeated. The \IterateLens combinator takes a lens as an
argument and synchronizes each element of the iteration with that lens. For
example, if \lstinline{employee_lens} synchronizes a single employee's data,
iterating that lens builds a lens that synchronizes a list of employees.
%
\begin{lstlisting}
iterate(id("\n") . employee_lens) : ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_ins)$^*$
\end{lstlisting}
%

\begin{figure}
\begin{lstlisting}
let name_lens = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
let employee_lens = name_lens . disc(salary,company,"N/A","UNK") in
let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
let emp_lens = or(id(""),employees_lens) in
ins(header) . employees_lens : emp_salaries $\Leftrightarrow$ emp_insurance
\end{lstlisting}
  \caption{A lens, written in our language for building simple symmetric lenses
    on string data, that synchronizes management and HR employee files.
    Note that we can annotate our lenses with types, to confirm the lens
    operates on the languages of the regular expressions. In this work, we
    synthesize the lens itself from such regular expressions.}
  \label{fig:example_lens}
\end{figure}

Using these combinators, we can build up a lens that synchronizes
management and HR employee files.

\ksf{This paragraph needs to be updated, if nothing else, to reflect
the new example. Also, we haven't mentioned any typing rules so the
first sentence doesn't make sense.  Perhaps the whole topic could be
defered to later where we can explain why such a rule is necessary?}
This lens uses one typing rule not previously
mentioned: type equivalence. While \lstinline{scheduled_job} is a equivalent to
a regular expression with an outermost disjunction, it is not syntactically
equivalent to one. Without type equivalence, the provided lens wouldn't actually
be well-typed, but type equivalence states that if a lens is typed between a
pair of regular expressions, it is also typed between an equivalent pair.

Note that, while the synchronization task is quite simple, the
requisite lens is relatively complex. Furthermore, this complexity is
inherent in manually writing simple symmetric lenses because the
lenses must precisely describe the portions of the data that should be
connected and those that should be disconnected.  For disconnected
pieces, the lens must further specifiy appropriate defaults.

@@

To alleviate these difficulties, we wish to synthesize symmetric lenses---given
a pair of regular expressions and a set of input-output examples, we want to
find a symmetric lens expression that is typed between the regular expression
pair, and satisfies the input/output examples. We call such a lens
a \emph{satisfying lens}.
In this example, we wish to
synthesize a lens from \lstinline{emp_salaries} and \lstinline{emp_insurance},
using the example data in Figure~\ref{fig:minimized-representations} as an
input-output example from which to draw the required defaut values for 
\Disconnect lenses.

A challenge is that the simple symmetric lens combinators permit many
well-typed lenses
between a given pair of regular expressions. For example, 
Figure~\ref{fig:example_lens} gives one possible lens between regular
expressions  \lstinline{emp_salaries} and \lstinline{emp_insurance},
but 
%
\begin{lstlisting}
disc(emp_salaries,emp_insurance,rep(emp_salaries),rep(emp_insurance))
\end{lstlisting}
%
is another valid lens (where \lstinline{rep($\Regex$)} is a string in the
language of $\Regex$). Furthermore, adding the synchronized files from
Figure~\ref{fig:minimized-representations} as an input-output example helps
little. The default strings given to \Disconnect merely need to be the strings
in the input-output examples, and the lens will still satisfy the specification.
While the disconnect lens is prohibited by two examples, or by a carefully
chosen input-output example on \PutR or \PutL, this problem continues presenting
itself for the sublenses, and is highlighted even more on larger
formats. 
\ksf{I don't understand the above.  We shouldn't explain the issue in
terms of \PutR and \PutL as we are focused on the combinators rather
than the raw functions.}
While
we could ask programmers to provide detailed and precise specifications for their
synthesis tasks, we would prefer the algorithm to choose the right lens
automatically. Instead of merely finding \emph{any} satisfying lens, we wish to
synthesize the \emph{best} satisfying lens.

But what is the best satisfying lens? We propose that the
best satisfying lens is the satisyfing lens that connects as much data
as possible between the two formats. More formally, we
propose that the best satisfying lens is the satisfying lens that 
minimizes the expected number of bits required to recover data in
one format from synchronized data in the other.  
In more detail, we say that two strings $s$ and $t$
are \emph{synchronized according to lens $l$} if 
$l.\PutR\ s\ t = t$ and $l.\PutL\ t\ s = s$ where the notation $l.\PutR$
extracts the \PutR function from simple symmetric lens $l$; similarly
for \PutL.
We say that we can \emph{recover} $s$ from $t$ using bits $b$ and lens
$l$ if we can reconstruct $s$ from $t$, $b$, and $l$.  For example,
given the \IdentityLens{} lens, we can recover $s$ from $t$ using
no bits because in this case, $s$ is just $t$.  In contrast, given
the \Disconnect{} lens, we need enough bits to fully encode $s$ in
order to recover it from $t$ because $t$ contains no useful
information.  This calculation shows that if both \IdentityLens{}
and \Disconnect{} are satisfying lenses, then \IdentityLens{} is
better.

The expected number of bits required to recover some data is a
well-known concept in information theory: it is the \emph{entropy} of
the data. Calculating the 
entropy of some data requires a probability distribution over the
space of possible values the data can take on.  Specifically,
given a set $S$ and a probability distribution over that set $P
\OfType S \rightarrow \Reals$, the entropy of the $(S,P)$ pair is $-\Sigma_{s\in
  S}P(s)\Log_2P(s)$.

In our setting, we already have a way of expressing sets of data:
regular expressions.  What we need is a way to express probability
distributions over that data.
To that end, we adopt \emph{stochastic regular expressions}\ksf{is
there a citation?}, 
which are regular expressions where each operator is annotated with a
probability.  Stochastic regular expressions
allow us to jointly specify the allowable data and give a probability
distribution over it.  From this information, we can calculate the
entropy of the stochastic regular expression.

Given a lens $\Lens$ between two stochastic regular expressions $\Regex$ and
$\RegexAlt$, we can calculate the expected number of bits required
to recover a string $s$ in $\LanguageOf{\Regex}$ from a synchronized
string $t$ in $\LanguageOf{\RegexAlt}$. 
This expected value is just the conditional entropy of the $\Regex$ given
$\RegexAlt$ and $\Lens$. 
To ensure symmetry, we seek to find the lens that
minimizes the sum of that entropy with the conditional entropy of $\RegexAlt$
given $\Regex$ and $\Lens$.

\ksf{This piece might better go somewhere deeper in the paper.}
Our algorithm guesses a reasonable probability distribution from
the syntax, though if users manually provided stochastic regular expressions, or
a large suite of data in the formats through which we could learn the annotations,
we could instead use that.

@@

Consider the trivial lens
\lstinline{disc(emp_salary,emp_row,rep(emp_salary),rep(emp_row))}. As this is
the disconnect lens, the cost of each side is the entropy of the data format, so
the total cost is $\EntropyOf{\lstinline{emp_salary}} +
\EntropyOf{\lstinline{emp_row}}$. \lstinline{emp_salary} contains strictly more
imformation than \EntropyOf{\lstinline{": " . salary}}, and \lstinline{emp_row}
contains strictly more information than \EntropyOf{\lstinline{": " . salary}},
so by our metric \lstinline{single_emp_map} is preferable to the disconnect
lens.
% end overview

@@

% begin implementation
\section{Implementation}
\label{sec:implementation}
We have implemented our symmetric DNF lenses and the synthesis of them as an
extension to Boomerang. We added symmetric lens combinators to Boomerang, and
implemented Boomerang's previous asymmetric lens combinators within the
symmetric lens combinators.

We have extended Optician with our symmetric synthesis algorithm.
% end implementation

% begin evaluation
\section{Evaluation}
\label{sec:evaluation}
% end evaluation

>>>>

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amsmath}
\usepackage{nccmath}
\usepackage{mathtools}
\usepackage{bussproofs}
\usepackage{varwidth}
\usepackage{amsthm}
\usepackage{csvsimple}
\usepackage{thmtools,thm-restate}
\usepackage{changepage}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{multirow,bigdelim}
\usepackage{multicol}
\usepackage{siunitx}
\usepackage{listings}
\usepackage{letltxmacro}
\usepackage{sansmath}
\usepackage{url}
\usepackage{flushend}
\usepackage{microtype}
\usepackage[utf8]{inputenc}
\usepackage{mathpartir}
\usepackage{empheq}
\usepackage{array}
\usepackage{pgfplots}
\usepackage{stmaryrd}
\usepackage{courier}
\usepackage{qtree}
\usepackage[normalem]{ulem}
\usepackage{relsize}
\usepackage{tikz}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{tabularx}
\usepackage{stackengine}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage{remreset}
\usepackage{tabulary}
\usepackage{xspace}

@@

\begin{figure}
  % https://tex.stackexchange.com/questions/168741/showing-two-listings-in-a-table-side-by-side
  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}[
stringstyle=\upshape\sffamily
]
Jane Doe: 35000
John Public: 37500
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingA}{\box0}

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}[
stringstyle=\upshape\sffamily
]
lastname,firstname,company
Doe,Jane,Healthcare Inc.
Public,John,Insurance Co.
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingB}{\box0}

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}
let salary = number | "UNSET"
let emp_salary = name . " " . name . ": " salary
let emp_salaries = "" | emp_salary . ("\n" emp_salary)*
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingC}{\box0}

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}
let company = (name . ("Co." | "Inc." | "Ltd.")) | "UNSET" in
let emp_row = name . "," . name "," company in
let header = "firstname,lastname,company" in
let emp_insurances = header . ("\n" . emp_row)*
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingD}{\box0}

  \centering
  \resizebox{\columnwidth}{!}{%
    \begin{tabular}{|>{\columncolor{lightyellow}}c|>{\columncolor{lightgreen}}c|}
      \hline
      company.mgmt & company.hr \\
      \hline
      \listingA & \listingB \\
      \hline
      \listingC & \listingD \\
      \hline
    \end{tabular}
  }
  \caption{Example files used by management and HR, and regular expressions
    describing the formats of these files.}
  \label{fig:minimized-representations}
\end{figure}

\afm{i commented out the previous version, because this one was quickly written
  up to give you something to work off. Feel free to delete it if it isnt
  helpful. } We motivate our problem with an extended example. In our
theoretical company, human resources and upper management both keep information
about employees. HR needs to know the names of the employees as well as their
insurance providers. Management needs to know the names of employees and their
salary. Two example files, and regular expressions describing the formats, are
shown in Figure~\ref{fig:minimized-representations}. These two files should be
synchronized: if management adds a new employee, ``Richard Roe: 32500'', a new
row should be added to HR's file, ``Roe,Richard,UNSET''. Furthermore, if HR
changes ``John Public'' to ``Jon Public'', that name change should be reflected
in both files. However, not all the data should be synchronized. The HR file
should not get altered due to salary changes, and the management file should not
get altered due to insurance company changes. While names should be matched
across the two formats, insurance company and salary information should be left
unmatched. Such a synchronization can be achieved with simple symmetric lenses.

@@

Lenses synchronize files by providing 4 functions, two \emph{create} functions,
and two \emph{put} functions. Create functions are used to initialize data in
one format when provided with data in the other. If a company.mgmt file is
initialized, a synchronized company.hr file should also be created, and vice
versa. Put functions are used to propagate edits from one format to the other.
If a company.hr file is updated, we should be able to combine it with a
non-updated company.mgmt file, and output a new company.mgmt file, synchronized
to the updated company.hr. To have a simple symmetric lens is to have these
required synchronization functions. Formally, a simple symmetric lens between
$X$ and $Y$ comprises 4 functions:
\begin{enumerate}
\item $\CreateR \OfType X \rightarrow Y$
\item $\CreateL \OfType Y \rightarrow X$
\item $\PutR \OfType X \rightarrow Y \rightarrow Y$
\item $\PutL \OfType Y \rightarrow X \rightarrow X$
\end{enumerate}
Furthermore, if a file is not edited, but is then resynchronized using a put
function, no data should be changed. To be a simple symmetric lens, one must
satisfy the following ``round-tripping'' laws, guaranteeing identity edits won't
change the file data.
\begin{gather}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{gather}

For synchronization, when the original company.hr or company.mgmt file is
created, a synchronized file of the other form is created from it with
$\CreateR$ or $\CreateL$ (depending on which file was initially created). On
future edits, changes are made with the puts, a \PutR would propagate the shared
information from the company.hr file, and knit it together with the information
specific to the company.mgmt file.

\subsection{Programming Symmetric Lenses} We now build up a symmetric lens by
example. Our symmetric lens expressions are typed between regular expression
pairs, if \lstinline{$\Lens$ : $\Regex \Leftrightarrow \RegexAlt$}, then $\Lens$
encodes a simple symmetric lens between $\LanguageOf{\Regex}$ and
$\LanguageOf{\RegexAlt}$. At the end, we want our symmetric lenses to by typed
between \lstinline{emp_salaries} and \lstinline{emp_insurance}.  The finished
lens is shown in Figure~\ref{fig:example_lens}.

The identity lens takes, as argument, a regular expression, and performs the
identity function on strings in that regular expression. The identity lens fully
matches the data formats: when one data format is updated, the other format gets
the exact same value. In our example, when a name is updated, it is also
expected to updated in the other format. The identity lens propagates that
information.
%
\begin{lstlisting}
id(name) : name $\Leftrightarrow$ name
\end{lstlisting}
%

Where the identity lens aligns data, the disconnect lens does the opposite: it
leaves data unaligned. The disconnect lens takes 4 arguments: two regular
expressions and two strings. On creates, because there is no association between
data in the first format and data in the second format, designated default
strings are used for the other format. In our extended example, salaries and
insurance companies should not be matched, so a disconnect lens would be used.
We want our default behavior to specify that the value has not been set.
%
\begin{lstlisting}
disc(salary,company,"UNSET","UNSET") : salary $\Leftrightarrow$ company
\end{lstlisting}
%
When building a disconnect lens between two strings, one of which is merely the
empty string, the lenses \lstinline{ins} and \lstinline{del} can instead be
used, where \lstinline{ins($t$) = disc("",$t$,"",$t$)} and
\lstinline{del($s$) = disc($s$,"",$s$,"")}.

Finally, there are a number of lenses that compositionally build up new lenses
from old ones. For example, data can appear in one form or another: a program
can run during a specific set of minutes, or that program can run the every
minute. If \lstinline{minutes_map} and \lstinline{alltimes_map} are lenses that
synchronize those settings in each format, then those lenses can be composed
with an \OrLens lens.
%
\begin{lstlisting}
or(id(""),multi_emp_map) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_row) . ("\n" emp_row)*
\end{lstlisting}

Data can also be concatenated, and so lens mappings can be concatenated.  This
is achieved with the concat map.
%
\begin{lstlisting}
concat(ins(header),multi_emp_map) : emp_salaries $\Leftrightarrow$ emp_insurance
\end{lstlisting}
%
Two lenses, $\Lens_1$ and $\Lens_2$, can be concatenated as
\lstinline{concat($\Lens_1$,$\Lens_2$)} or as the infix style
\lstinline{$\Lens_1$ . $\Lens_2$}.

Lastly, the \IterateLens lens synchronizes repeated data, aligning the number of
times the data is repeated. The \IterateLens combinator takes a lens as an
argument and synchronizes each element of the iteration with that lens. For
example, if \lstinline{single_emp_map} synchronizes a single employee's data,
iterating that lens builds a lens that synchronizes a list of tasks.
%
\begin{lstlisting}
iterate(id("\n") . single_emp_map) : ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_row)$^*$
\end{lstlisting}
%

\begin{figure}
\begin{lstlisting}
let name_map = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
let single_emp_map = name_map . disc(salary,company,"UNSET","UNSET") in
let multi_emp_map = ins("\n") . single_emp_map . iterate(id("\n") . single_emp_map) in
let emp_map = or(id(""),multi_emp_map) in
ins(header) . multi_emp_map : emp_salaries $\Leftrightarrow$ emp_insurance
\end{lstlisting}
  \caption{A lens, written in our language for building simple symmetric lenses
    on string data, that synchronizes management and hr employee files.
    Note that we can annotate our lenses with types, to confirm the lens
    operates on the languages of the regular expressions. In this work, we
    synthesize the lens itself from such regular expressions.}
  \label{fig:example_lens}
\end{figure}

Using these combinators, we can build up a lens that synchronizes HR and
management employee files. This lens uses one typing rule not previously
mentioned: type equivalence. While \lstinline{scheduled_job} is a equivalent to
a regular expression with an outermost disjunction, it is not syntactically
equivalent to one. Without type equivalence, the provided lens wouldn't actually
be well-typed, but type equivalence states that if a lens is typed between a
pair of regular expressions, it is also typed between an equivalent pair.

Note that, while the programmed lens should be quite simple, it was relatively
difficult to write. We had to provide a lot of scaffolding to map between the
desired formats, even though the idea was quite simple: align the times but
leave the commands unaligned. Furthermore, this is inherent to manually writing
these synchronization functions as a programmer must describe which components
of the strings are aligned with which, and which components are scaffolding.

@@

To alleviate these difficulties, we wish to synthesize symmetric lenses---given
a pair of regular expressions and a set of input-output examples, we want to
find a symmetric lens expression that is typed between the regular expression
pair, and satisfies the input/output examples. In this example, we wish to
synthesize a lens from \lstinline{emp_salaries} and \lstinline{emp_insurance},
using the strings in Figure~\ref{fig:minimized-representations} as an
input-output example for \CreateR.

However, our simple symmetric lens combinators permit many well-typed programs
between two regular expressions. For example, one possible lens between
\lstinline{emp_salaries} and \lstinline{emp_insurance} is the one in
Figure~\ref{fig:example_lens}, but the lens
%
\begin{lstlisting}
disc(emp_salaries,emp_insurance,rep(emp_salaries),rep(emp_insurance))
\end{lstlisting}
%
is also a valid lens (where \lstinline{rep($\Regex$)} is a string in the
language of $\Regex$). Furthermore, adding the synchronized files from
Figure~\ref{fig:minimized-representations} as an input-output example helps
little. The default strings given to \Disconnect merely need to be the strings
in the input-output examples, and the lens will still satisfy the specification.
While the disconnect lens is prohibited by two examples, or by a carefully
chosen input-output example on \PutR or \PutL, this problem continues presenting
itself for the sublenses, and is highlighted even more on larger formats. While
we could merely ask programmers to provide large, tight specifications for their
synthesis tasks, we would prefer the algorithm to choose the right lens
automatically. Instead of merely finding \emph{any} satisfying lens, we wish to
synthesize the \emph{best} satisfying lens.

But what is the best satisfying lens? Here, we take the view that the best satisfying lens is the
one that aligns as much data as is permitted by the type system and the
examples. More formally, we wish to minimize the expected number of bits
required to recover data in one format from a synchronized string in the other.
The expected number of bits is a well known concept in information theory: the
\emph{entropy} of that data source.

The entropy of a data source is based on the probability distribution of that
data source. Given a set $S$, and a probability distribution over that set $P
\OfType S \rightarrow \Reals$, the entropy of the $(S,P)$ pair is $-\Sigma_{s\in
  S}P(s)\Log_2P(s)$. While we have ways of expressing sets of strings (with
regular expressions), we do not have probability distributions. Here, we
express these probability distributions jointly with the language through
\emph{stochastic regular expressions}, regular expressions annotated with
probabilities. Our algorithm guesses a reasonable probability distribution from
the syntax, though if users manually provided stochastic regular expressions, or
a large suite of data in the formats through which we could learn the annotations,
we could instead use that.

With sets and probability distributions, we can find the expected number of bits
required to recover one string from a synchronized string. This expected value
is the conditional entropy of the stochastic regular expression ($\Regex$) given
another stochastic regular expression ($\RegexAlt$), and a lens that shows which
strings are synchronized with which ($\Lens$). We aim to find the lens that
minimizes the sum of that entropy with the conditional entropy of $\RegexAlt$
given $\Regex$ and $\Lens$.

@@

Consider the trivial lens
\lstinline{disc(emp_salary,emp_row,rep(emp_salary),rep(emp_row))}. As this is
the disconnect lens, the cost of each side is the entropy of the data format, so
the total cost is $\EntropyOf{\lstinline{emp_salary} } +
\EntropyOf{\lstinline{emp_row}}$. \lstinline{emp_salary} contains strictly more
imformation than $\EntropyOf{\lstinline{": " . salary}}$, and
\lstinline{emp_row} contains strictly more information than
\EntropyOf{\lstinline{": " . salary}}, so by our metric
\lstinline{single_emp_map} is preferable to the disconnect lens.
% end overview

@@

% begin implementation
\section{Implementation}
\label{sec:implementation}
% end implementation

% begin evaluation
\section{Evaluation}
\label{sec:evaluation}
We have implemented our symmetric DNF lenses and the synthesis of them as an
extension to Boomerang. We added symmetric lens combinators to Boomerang, and
implemented Boomerang's previous asymmetric lens combinators within the
symmetric lens combinators.

We have extended Optician with our symmetric synthesis algorithm. Users can
write synthesis tasks alongside lens combinators, use those synthesis results in
the manual writing of new lenses, and use previously defined lenses in the
synthesis of new lenses.

In this work, we aim to answer two primary research questions:
\begin{itemize}
\item Is our synthesis procedure efficient enough to be used in everyday
  development?  Can it be used in place of the existing bijective lens synthesis
  algorithm integrated with Boomerang?
\item Is our information theoretic cost metric useful in finding good lenses.
\end{itemize}

\subsection{Benchmark Suite Construction}
Our benchmarks came from four different sources.
\begin{enumerate}
\item We adapted \afm{8} data cleaning benchmarks from Flash Fill. In adopting
  these benchmarks, we go beyond the data cleaning tasks, and synthesize
  synchronization functions between the ad-hoc data and their cleaned forms.
  Optician also adopted benchmarks from Flash Fill, but required alterations to
  be in bijective form. With our symmetric enhancement, we Optician's four
  benchmarks, and 4 new Flash Fill benchmarks with no alterations.
\item We adapted \afm{29} benchmarks from Augeas.  These benchmarks are the same
  as those present in Optician.  While these benchmarks had to be slightly
  adopted for bijectivity, previous work addressed this lack of bijectivity by
  synthesizing quotient bijective lenses~\cite{?}.
\item We used \afm{8} benchmarks from synchronization tasks on synchronizing a
  variety of real-life formats. These formats range from the synchronizing REST
  and JSON web resource descriptions to synchronizing BibTeX and EndNote
  citation descriptions.
\end{enumerate}

\subsection{Efficiency}

To answer our first research question, we analyze the runtime of our algorithm,
and compare it to the existing algorithm for synthesizing bijective lenses.

We synthesized lenses in 4 modes:

\begin{tabulary}{\linewidth}{rL}
  \SSOpt{}: & We synthesize lenses using our symmetric synthesis algorithm with all optimizations enabled.\\
  \SSNCOpt{}: & We synthesize lenses using our symmetric synthesis algorithm,
                but compositional synthesis is not enabled.\\
  \BSOpt{}: &  We synthesize lenses using the existing bijective lens synthesis
              algorithm with all optimizations enabled\\
  \BSNCOpt{}: & We synthesize lenses using the existing bijective lens synthesis
                algorithm, but compositional synthesis is not enabled.\\
\end{tabulary}\\

We include running these algorithms without compositional synthesis to stress
test the algorithms. While compositional synthesis makes the procedure easier,
and demonstrates the potential to be used in arbitrarily hard problems, it is
dependent on the user identifying important sublenses between subcomponents of
the final formats. It is therefore interesting to see how many problems each
algorithm can solve, when trying to synthesize ``in one go''.

For each benchmark in our benchmark suite, we ran every mode on it with a
timeout of 1 minute, and average the result over 5 runs. Figure~\ref{fig:times}
summarizes the results of these tests.
% end evaluation
